### 2025-03-23 plots/elision/mem_summary.tex
Flesh out summaries for experiments in section 9

\begin{tabular}{lrr}
\toprule
configuration & elision-naive & elision-opt \\
suite &  &  \\
\midrule
\binarytrees & 1.06MiB \footnotesize{± 22.22KiB} & 1.06MiB \footnotesize{± 24.18KiB} \\
\regexredux & 74.37MiB \footnotesize{± 3.15MiB} & 72.76MiB \footnotesize{± 3.78MiB} \\
\somrsast & 1.16MiB \footnotesize{± 517.62KiB} & 1.22MiB \footnotesize{± 1.20MiB} \\
\somrsbc & 763.30KiB \footnotesize{± 312.94KiB} & 692.14KiB \footnotesize{± 370.14KiB} \\
\yksom & 926.88KiB \footnotesize{± 189.68KiB} & 1.37MiB \footnotesize{± 160.51KiB} \\
\bottomrule
\end{tabular}