### 2025-03-23 plots/premopt/mem_summary.tex

\begin{tabular}{lrrr}
\toprule
configuration & premopt-naive & premopt-none & premopt-opt \\
suite &  &  &  \\
\midrule
\binarytrees & 1.06MiB \footnotesize{± 18.74KiB} & 1.06MiB \footnotesize{± 16.18KiB} & 1.06MiB \footnotesize{± 26.75KiB} \\
\regexredux & 73.57MiB \footnotesize{± 3.46MiB} & 72.60MiB \footnotesize{± 3.05MiB} & 73.25MiB \footnotesize{± 3.35MiB} \\
\somrsast & 1.22MiB \footnotesize{± 1.21MiB} & 1.23MiB \footnotesize{± 1.08MiB} & 1.22MiB \footnotesize{± 1.15MiB} \\
\somrsbc & 694.50KiB \footnotesize{± 347.61KiB} & 689.14KiB \footnotesize{± 330.27KiB} & 701.91KiB \footnotesize{± 372.72KiB} \\
\yksom & 1.38MiB \footnotesize{± 165.17KiB} & 1.37MiB \footnotesize{± 161.23KiB} & 1.37MiB \footnotesize{± 153.53KiB} \\
\bottomrule
\end{tabular}