### 2025-03-26 plots/elision/mem_summary.tex
Put the correct elision mem data in table 6

\begin{tabular}{lrrrrrrr}
\toprule
configuration & Elision Naive & Elision Opt \\
suite &  &  \\
\midrule
\binarytrees & 1.07MiB \footnotesize{(1.05MiB-1.09MiB)} & 1.06MiB \footnotesize{(1.04MiB-1.09MiB)} \\
\regexredux & 75.35MiB \footnotesize{(72.13MiB-78.58MiB)} & 74.19MiB \footnotesize{(70.88MiB-77.50MiB)} \\
\midrule
\alacritty & 8.69MiB \footnotesize{(8.23MiB-9.18MiB)} & 8.65MiB \footnotesize{(8.20MiB-9.14MiB)} \\
\fd & 3.51MiB \footnotesize{(3.50MiB-3.53MiB)} & 3.51MiB \footnotesize{(3.49MiB-3.52MiB)} \\
\grmtools & 11.71MiB \footnotesize{(11.58MiB-11.84MiB)} & 60.08MiB \footnotesize{(59.42MiB-60.75MiB)} \\
\ripgrep & 5.22MiB \footnotesize{(5.19MiB-5.24MiB)} & 7.71MiB \footnotesize{(7.66MiB-7.76MiB)} \\
\midrule
\somrsast & 1.14MiB \footnotesize{(1.14MiB-1.15MiB)} & 1.08MiB \footnotesize{(1.08MiB-1.09MiB)} \\
\somrsbc & 740.48KiB \footnotesize{(736.99KiB-743.99KiB)} & 629.92KiB \footnotesize{(624.19KiB-635.70KiB)} \\
\yksom & 931.19KiB \footnotesize{(927.32KiB-935.07KiB)} & 1.28MiB \footnotesize{(1.27MiB-1.29MiB)} \\
\bottomrule
\end{tabular}