### 2025-07-18 tables/elision_finalizer_counts.tex
Restructure Evaluation section and add additional plots

\begin{tabular}{l r@{\hspace{3pt}}l r@{\hspace{3pt}}l r@{\hspace{3pt}}l}
\toprule
Suite & \multicolumn{2}{c}{Time in GC} & \multicolumn{2}{c}{Wall-clock} & \multicolumn{2}{c}{User-time} \\
\midrule
alacritty & 1.01$\times$ & \scriptsize\textcolor{gray!60}{[0.90,1.14]} & 1.01$\times$ & \scriptsize\textcolor{gray!60}{[0.84,1.24]} & 1.00$\times$ & \scriptsize\textcolor{gray!60}{[0.74,1.36]} \\
fd & 1.01$\times$ & \scriptsize\textcolor{gray!60}{[0.75,1.38]} & 1.02$\times$ & \scriptsize\textcolor{gray!60}{[0.75,1.47]} & 0.97$\times$ & \scriptsize\textcolor{gray!60}{[0.53,1.62]} \\
grmtools & 0.14$\times$ & \scriptsize\textcolor{gray!60}{[0.11,0.17]} & 0.67$\times$ & \scriptsize\textcolor{gray!60}{[0.64,0.69]} & 0.54$\times$ & \scriptsize\textcolor{gray!60}{[0.52,0.56]} \\
ripgrep & 0.47$\times$ & \scriptsize\textcolor{gray!60}{[0.43,0.56]} & 0.95$\times$ & \scriptsize\textcolor{gray!60}{[0.87,1.03]} & 0.87$\times$ & \scriptsize\textcolor{gray!60}{[0.75,1.00]} \\
som-rs-ast & 0.13$\times$ & \scriptsize\textcolor{gray!60}{[0.08,0.23]} & 0.13$\times$ & \scriptsize\textcolor{gray!60}{[0.12,0.15]} & 0.09$\times$ & \scriptsize\textcolor{gray!60}{[0.08,0.10]} \\
som-rs-bc & 0.08$\times$ & \scriptsize\textcolor{gray!60}{[0.05,0.14]} & 0.09$\times$ & \scriptsize\textcolor{gray!60}{[0.09,0.10]} & 0.07$\times$ & \scriptsize\textcolor{gray!60}{[0.06,0.07]} \\
\midrule
overall & 0.98$\times$ & \scriptsize\textcolor{gray!60}{[0.75,1.31]} & 0.26$\times$ & \scriptsize\textcolor{gray!60}{[0.24,0.28]} & 0.20$\times$ & \scriptsize\textcolor{gray!60}{[0.18,0.23]} \\
\bottomrule
\end{tabular}