### 2025-07-28 response.tex
Add some detailed comments to the response

\documentclass[12pt,a4paper,preprint]{article}

\usepackage{setspace}
\usepackage{xcolor}
\usepackage{amssymb}
\usepackage{listings}
\usepackage{listings-rust}
\usepackage{softdev}
\usepackage{quoting}
\usepackage{xspace}
\usepackage[many]{tcolorbox}
\newtcolorbox{blockquote}{%
  sharp corners,
  colback=gray!5,
  colframe=gray!60,
  leftrule=2pt,
  enhanced,
  before skip=6pt, after skip=18pt,
  boxrule=0pt,
  left=6pt, right=4pt,
  borderline west={2pt}{0pt}{gray!60},
  fontupper=\scriptsize\sffamily\selectfont
}

\setlength{\parindent}{0pt}
\setlength{\parskip}{2.0ex plus0.5ex minus0.2ex}

%\newcommand\laurie[1]{\mynote{Laurie}{#1}}
%\newcommand\jake[1]{\mynote{Jake}{#1}}

\newcommand\Egcrc{E$_\textrm{gcvs}$\xspace}
\newcommand\Eelision{E$_\textrm{elision}$\xspace}
\newcommand\Epremopt{E$_\textrm{premopt}$\xspace}
\newcommand\boehm{\textsc{BDWGC}\xspace}

\begin{document}

\date{}  % This removes the date
\title{Response to Reviewer Feedback}
\maketitle

We thank the reviewers for their thoughtful reviews: we hope to have
taken onboard your feedback and addressed your questions. In this document
we summarise the major changes we have made: the diff shows changes to the prose
(though it does tend to mangle changes to figures and tables).

\subsection*{Finalization semantics}

\begin{blockquote}
explicitly discuss and present the issues around finalizations semantics, as
  learned from other languages
\end{blockquote}

We have made a number of changes (chiefly to the end of Section 3, and
throughout Section 4), which we hope address this.

As the reviewers note ``an
implementation that does not finalize anything would be compliant when
finalization may be deferred indefinitely'' is true of most GCed
implementations we are aware of, including Java. Indeed, Java once had the
\lstinline{runFinalizersOnExit} flag to force finalizers to be run before the
JVM exists, but it is marked as deprecated as ``inherently unsafe''; instead,
there is now the \lstinline{addShutdownHook} mechanism.

While investigating this further, we discovered one surprise (to us, at least),
which muddies the waters with respect to finalizer latency further. In Rust,
when a thread unwinds (i.e.~panics) then, depending on compile-time flags, no
destructors will be run --- but Alloy can run all of the `queued' finalizers!
We have not belaboured this point in the paper, but it shows that destructors
can have guaranteed-infinite latency in some situations where finalizers have
infinite-at-worst latency.

\subsection*{Performance Evaluation}

\jake{Since the original submission, alloy has been updated. We now
dynamically link boehm in all cases. For memory metrics, this alone won't
change things because we always used dynamic linkage for these tests anyway.
However, this may have affected wall-clock performance on some individual
benchmarks (though it's really hard to tell which, because there are
confounders: we also now run benchmarks on b15; we made changes to finalization
based on Andy wingo's feedback (mentioned below); and we made a couple of other
unrelated QoL and implementation changes to Alloy (i doubt these made a
difference, but who knows). What is clear though is that even if individual
benchmarks may have changed, the overall picture has not not since original
submission.}

We have made a large number of changes to the performance evaluation. These are
mostly due to reviewer comments, but we also uncovered a mistake in our
experiment. We used \boehm's \lstinline{GC_heapsize} metric (which never
shrinks) for calculating in-use heap snapshots, we have since modified boehm to
output the number of bytes swept and explicitly freed on each GC. While neither mistake
fundamentally changes the outcome, we believe it is important that we not
try and sweep this under the carpet. \jake{This does fundamentally change the
memory results for finalizer elision. In som-rs-ast, som-rs-bc, grmtools, and
ripgrep, finalizer elision now uses less memory than the no-elision
configuration. In the original submission, we incorrectly observed that it had
a higher average heap usage as a trade-off for better wall-clock performance.}
\jake{furthermore, since the original submission we have changed the way that
finalization works: we now use Boehm's \lstinline{finalize_on_demand} API based on
Andy Wingo's feedback. It is very possible that this could have moved the dial
on heap size too, how much that this changes things in relation to the above
mentioned wrong metric error I don't know. but it has crossed my mind.}

The most significant changes are:

\begin{itemize}
  \item We provide a comparison of \lstinline{Rc<T>} with jemalloc and \boehm
    in Table~\laurie{3?} in the Appendix. This suggests that the inherent
    overhead of \boehm ranges from 2--26\%, averaging around 11\%.

  \item We have standardized all visualizations by normalizing against
    appropriate baselines (e.g.~Alloy-without-elision as the baseline for
    the \Eelision experiment; and barrier-free execution as the baseline
    for the \Epremopt experiment). This allows us to show what we hope
    are easily understand performance ratios. We also provide the raw data
    in the appendix.

  \item We provide data for wall-clock and user-time.

  \item We provide data for the number of collections run, and the total GC
  pause time.
\end{itemize}

The downside to this is that we now present far too many figures and tables to
fit in the main body of the paper! We have tried to select the subset of
visualizations that will most quickly help the reader understand the overall
picture for the main body of the paper, relegating everything else to the
appendix.

We were asked -- when explaining the overhead of \boehm -- the following:

\begin{blockquote}
Ideally this would include a
discussion of how much of the performance deficit is likely redeemable via
"engineering", and how much is likely to be intrinsic (i.e. due to
unavoidable conservatism), so that we can get a better intuition of the
plausible "best case" for this approach.
\end{blockquote}

Although we can now quantify the overhead of Boehm vs.~jemalloc in our context,
the data doesn't really give us clues as to why the overhead exists. That makes
us nervous about speculating further. While we are not total allocator dummies,
both of us have talked to allocator experts over the years to know that there is a lot
here that we don't know. At a high level, \boehm is -- and we don't mean this as an
insult or criticism! -- ``old''. Since then many allocators (e.g.~dlmalloc, jemalloc)
have been designed, and subsequently tweaked and improved; each such allocator
is full of many decisions which are likely to affect its performance. Unfortunately,
we don't really know which decisions might be relevant or not.

\jake{My reading of this reviewer comment is that our mention of allocator
differences are just a partial (yet important!) answer. I think they are also
asking something else too (based on their mention of ``unavoidable
conservatism''). This is more along the lines of: ``what GC-specific
engineering would be needed to make this research-based GC productionised to
have comparable performance with jemalloc RC''. My intuition is that they're
wondering about things like ``what if you had a Bartlett-style
semi-conservative collector rather than fully conservative?'' The short answer
here would also be ``it's too difficult to say'' IMHO. Bartlett-style
collection muddies the waters a bit showing that (counter to what I would
assume) precise object scanning doesn't make a lot of difference in terms of
wall-clock time. For some workloads its slower, for others its a modest
improvement. And trying to make guesses on Rust workloads based off of this
would be pure speculation on my part. Precise GC OTOH would be an interesting
comparison, maybe fully moving compaction would improve things... but as we
mention in the paper, this is not possible while respecting the Rust semantics
of today. If you think this is worth mentioning in the body of the paper I can
do so and include a citation to the Bartlett work. I don't know how on-topic it
is though.}

\subsection*{Conclusion}

\begin{blockquote}
add a discussion thats sums up the overall benefits/drawbacks of this
particular approach, possibly depending on specific scenarios/use cases, and
perhaps outlines directions for future work from a conceptional perspective,
i.e., possible design directions that could be investigated
\end{blockquote}

We have expanded the conclusions, but we hope to have done so in a way
that does not overstate our confidence in our predictions.

\end{document}