### 2025-07-29 tables/exclusions.tex
Flesh out excluded benchmark appendix table

\begin{tabular}{lll}
\toprule
Benchmark         & Description                              & Reason for exclusion                                       \\
\midrule
bevy          & ECS game engine in Rust                   & \jake{TODO} \\
dyon          & Scripting language in Rust                & \jake{TODO} \\
jiff          & A datetime library for Rust               & \jake{TODO} \\
mini-moka     & Concurrent in-memory cache library        & \jake{TODO} \\
numbat        & Math search engine                        & \jake{TODO} \\
rkyv          & Zero-copy deserialization framework       & \jake{TODO} \\
RustPython    & Python interpreter written in Rust        & \jake{TODO} \\
rust-analyzer & Language server for Rust                  & \jake{TODO} \\
salsa         & Incremental recomputation library         & \jake{TODO} \\
WLambda       & scripting language in Rust                & \jake{TODO} \\
\bottomrule
\end{tabular}