### 2025-07-29 tables/exclusions.tex
Update benchmark exclusions table in appendix

\begin{tabular}{lll}
\toprule
Benchmark         & Description                              & Reason for exclusion                                       \\
\midrule
bevy          & ECS game engine in Rust                   & Unable to port successfully (see \cref{rustpython}) \\
dyon          & Scripting language in Rust                & Unable to port successfully (see \cref{rustpython})\\
jiff          & A datetime library for Rust               & Too few allocations to measure \\
mini-moka     & Concurrent in-memory cache library        & Too few allocations to measure \\
numbat        & Math search engine                        & Too few allocations to measure \\
rkyv          & Zero-copy deserialization framework       & Insufficient \texttt{Gc<T>} coverage in benchmarks \\
RustPython    & Python interpreter written in Rust        & Difficulty retro-fitting \texttt{\_\_del\_\_} semantics (see \cref{rustpython}). \\
rust-analyzer & Language server for Rust                  & Unable to port successfully (see \cref{rustpython}) \\
salsa         & Incremental recomputation library         & Too few allocations to measure \\
WLambda       & Scripting language written in Rust        & Insufficient \texttt{Gc<T>} coverage in benchmarks \\
\bottomrule
\end{tabular}