### 2025-07-30 tables/gcvs_perf_summary.tex
Remove all red and fix all todos

\begin{tabular}{l l r@{\hspace{3pt}}l r@{\hspace{3pt}}l}
\toprule
Suite & Allocator & Wallclock time (s) & & Slowdown & \\
\toprule
alacritty & \texttt{jemalloc}       & 0.39 & \scriptsize\textcolor{gray!60}{[0.36,0.43]} & -           &  \\
          & \texttt{boehm}          & 0.40 & \scriptsize\textcolor{gray!60}{[0.38,0.43]} & 1.02$\times$ & \scriptsize\textcolor{gray!60}{[0.87,1.21]} \\[2pt]
fd        & \texttt{jemalloc}       & 0.28 & \scriptsize\textcolor{gray!60}{[0.25,0.34]} & -           &  \\
          & \texttt{boehm}          & 0.31 & \scriptsize\textcolor{gray!60}{[0.27,0.35]} & 1.08$\times$ & \scriptsize\textcolor{gray!60}{[0.80,1.41]} \\[2pt]
grmtools  & \texttt{jemalloc}       & 2.89 & \scriptsize\textcolor{gray!60}{[2.82,2.97]} & -           &  \\
          & \texttt{boehm}          & 3.10 & \scriptsize\textcolor{gray!60}{[3.01,3.18]} & 1.07$\times$ & \scriptsize\textcolor{gray!60}{[1.01,1.13]} \\[2pt]
ripgrep   & \texttt{jemalloc}       & 1.43 & \scriptsize\textcolor{gray!60}{[1.36,1.51]} & -           &  \\
          & \texttt{boehm}          & 1.46 & \scriptsize\textcolor{gray!60}{[1.39,1.53]} & 1.02$\times$ & \scriptsize\textcolor{gray!60}{[0.92,1.12]} \\[2pt]
som-rs-ast & \texttt{jemalloc}      & 0.58 & \scriptsize\textcolor{gray!60}{[0.56,0.60]} & -           &  \\
           & \texttt{boehm}         & 0.73 & \scriptsize\textcolor{gray!60}{[0.70,0.75]} & 1.26$\times$ & \scriptsize\textcolor{gray!60}{[1.15,1.34]} \\[2pt]
som-rs-bc  & \texttt{jemalloc}      & 0.23 & \scriptsize\textcolor{gray!60}{[0.22,0.24]} & -           &  \\
           & \texttt{boehm}         & 0.27 & \scriptsize\textcolor{gray!60}{[0.26,0.29]} & 1.18$\times$ & \scriptsize\textcolor{gray!60}{[1.10,1.28]} \\[2pt]
overall   & \texttt{jemalloc}       & 0.49 & \scriptsize\textcolor{gray!60}{[0.47,0.51]} & -           &  \\
          & \texttt{boehm}          & 0.56 & \scriptsize\textcolor{gray!60}{[0.53,0.58]} & 1.14$\times$ & \scriptsize\textcolor{gray!60}{[1.05,1.24]} \\
\bottomrule
\end{tabular}